A weakly structured stem for human origins in Africa

### 2023-02-26 paper.tex
Save condensed abstract

\begin{abstract}
Despite broad agreement that \emph{Homo sapiens}
originated within Africa, considerable
uncertainty surrounds specific models of divergence and migration across the
continent. Progress is hampered by limited fossil and genomic data, as
well as variability in prior divergence time estimates.
We use linkage disequilibrium and
diversity-based statistics, optimized for rapid, complex demographic inference
to discriminate among such models.
We infer detailed demographic models for
populations across Africa, including representatives from eastern and western
groups, as well as new whole-genome sequence data for 44 Nama (Khoe-San) individuals. 
We infer a reticulated African population history in which contemporary
population structure dates back to Marine Isotope Stage (MIS) 5. The earliest
population divergence among contemporary populations occurs 120-135ka, between
the Khoe-San and other groups. Prior to the divergence of contemporary African
groups, we infer structure between two or more weakly
differentiated ancestral \emph{Homo}
populations connected by gene flow over
hundreds of thousands of years (i.e., a weakly structured stem). Weakly structured stem 
models provide more likely explanations of patterns of polymorphism
that had previously been attributed to contributions from archaic hominins in
Africa.
In contrast to models with archaic introgression, we predict that fossil
remains from coexisting ancestral populations should be morphologically
similar.
Despite genetic similarity between these populations, an inferred 1--4\%
of genetic differentiation among contemporary human populations can be
attributed to genetic drift between stem populations.
We show that model misspecification explains variation in previous
divergence time estimates and argue that studying a suite of models is key to
robust inferences about deep history.
\end{abstract}

\begin{abstract}
    \aprcomment{An attempt at condensing a bit -- this is just under 230 words}

>>>>

\begin{abstract}
